---
title: "ANA 515 Visualizations"
author: "Amanda Thiell"
date: "4/17/2022"
output:
  html_document:
    theme:
      bootswatch: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install, include=FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/"))
install.packages("ggplot2")
library(ggplot2)
```

```{r dataset, include=FALSE}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv"
recentgrads <- read.csv(url)
```

The first visualization seen below explores the variance in unemployment rates per major, split by the associated major category in a series of scatterplots. This arrangement allows easy evaluation of which major category showed results of the lowest and highest rates of unemployment.

```{r plot, include=TRUE}
ggplot(recentgrads, aes(Major, Unemployment_rate)) + geom_point(aes(color = Major_category)) + scale_y_continuous("Unemployment Rate", breaks = seq(0,0.5,by = 0.05))+ theme_bw() + labs(title="Scatterplot - Unemployment Rate by Major Category") + facet_wrap( ~ Major_category)
```

Second, we can review the total number of individuals in each major category, which includes all associated majors for each. As shown below, there is a large gap between the number of students in a "business" degree compared to other categories. In the distribution below, major "food science" contained no data, so that was removed from the comparison.

```{r totals, include=TRUE}
ggplot(recentgrads, aes(Major_category, Total)) + geom_bar(stat = "identity", fill = "magenta") + scale_x_discrete("Major Category")+ scale_y_continuous("Number of Individuals", breaks = seq(0,1600000, by = 100000))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + labs(title = "Bar Chart - Individuals per Major Category")
```

